---
title: Exact likelihood formula for the linear birth-death process with destructive sampling
author: A.A. King
output: rmarkdown::html_vignette
bibliography: phylopomp.bib
nocite: |
  @King2024, @King2022, @Stadler2010, @Gavryushkina2014
csl: jss.csl
vignette: >
  %\VignetteIndexEntry{Exact likelihood for LBDP with destructive sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_setup}
#| cache: false
#| purl: false
#| include: false
library(knitr)
opts_chunk$set(
             progress=TRUE,
             prompt=FALSE,
             tidy=FALSE,
             highlight=TRUE,
             strip.white=TRUE,
             message=FALSE,
             warning=FALSE,
             error=TRUE,
             echo=TRUE,
             cache=TRUE,
             cache.extra=rand_seed,
             results='markup',
             fig.show='asis',
             fig.path=paste0("tmp/","lbdp_destruc","/"),
             cache.path=paste0("tmp/","lbdp_destruc","/"),
             size='small',
             fig.align='center',
             fig.dim=c(6,6),
             dpi=100,
             dev='png',
             dev.args=list(bg='transparent')
           )
options(
  keep.source=TRUE,
  encoding="UTF-8"
)
```

```{r packages}
#| cache: false
#| include: false
library(tidyverse)
library(pomp)
library(phylopomp)
library(cowplot)
stopifnot(getRversion() >= "4.5")
stopifnot(packageVersion("pomp")>="6.2")
stopifnot(packageVersion("phylopomp")>="0.16.2")
theme_set(theme_bw(base_family="serif"))
options(
  dplyr.summarise.inform=FALSE,
  pomp_archive_dir="results/lbdp_destruc"
)
set.seed(1159254136)
```

The following codes simulate a realization of the genealogy process induced by a linear birth-death-sampling process with birth rate $\lambda$, death rate $\mu$, non-destructive sampling rate $\psi$, and destructive sampling rate $\chi$.


```{r lbdp1}
true.params <- data.frame(lambda=3.5,mu=0.8,psi=1,chi=2,n0=20)
true.params |>
  with(
    runLBDP(lambda=lambda,mu=mu,psi=psi,chi=chi,n0=n0,time=3)
  ) -> x
```

```{r lbdp1_plot}
#| echo: false
plot_grid(
  x |> plot(points=TRUE),
  x |> lineages() |> plot(),
  ncol=1,
  align="hv",
  rel_heights=c(2,1)
)
```

The following designs a slice through parameter space in the direction of the parameter $r=\chi/(\psi+\chi)$, which is the fraction of the samples that are destructive.

```{r lbdp2}
expand_grid(
  lambda=3.5,
  mu=0.8,
  r=seq(0,1,by=0.01),
  totsamp=with(true.params,psi+chi),
  n0=20
  ) |>
  mutate(
    psi=totsamp*(1-r),
    chi=totsamp*r
  ) -> params
```

Next, we perform a sequential Monte Carlo (particle filter) computation at each point in the above design.

```{r lbdp3}
library(doParallel)
library(iterators)
foreach (
  p=iter(params,"row"),
  .inorder=FALSE,
  .combine=bind_rows
) %dopar% {
  p |>
    with(
      x |>
        lbdp_exact(lambda=lambda,mu=mu,psi=psi,chi=chi,n0=n0)
    ) -> ll1
  bind_cols(p,exact=ll1)
} |>
  bake(file="slice1.rds") -> params
```

```{r lbdp3_plot}
#| echo: false
params |>
  filter(is.finite(exact)) |>
  ggplot(aes(x=r,y=exact))+
  geom_line()+
  geom_vline(xintercept=with(true.params,chi/(psi+chi)),linetype=2)+
  labs(
    color=character(0),
    y="log likelihood",
    x=expression(r)
  )+
  theme(
    legend.position="inside",
    legend.position.inside=c(0.5,0.22),
    legend.background=element_rect(fill="white")
  )
```

```{r lbdp4a}
expand_grid(
  lambda=3.5,
  mu=0.8,
  psi=exp(seq(log(0.1),log(8),length=100)),
  chi=exp(seq(log(0.1),log(8),length=100)),
  n0=20
) -> params
```

```{r lbdp4b}
library(doParallel)
library(iterators)
registerDoParallel(cores=20)
foreach (
  p=iter(params,"row"),
  .inorder=FALSE,
  .combine=bind_rows
) %dopar% {
  p |>
    with(
      x |>
        lbdp_exact(lambda=lambda,mu=mu,psi=psi,chi=chi,n0=n0)
    ) -> ll1
  bind_cols(p,exact=ll1)
} |>
  bake(file="slice2.rds",dependson=params) -> params
```

```{r lbdp4_plot}
#| echo: false
params |>
  filter(is.finite(exact)) |>
  ggplot(aes(x=psi,y=chi,z=exact))+
  geom_contour(binwidth=30)+
  geom_vline(xintercept=with(true.params,psi),linetype=2)+
  geom_hline(yintercept=with(true.params,chi),linetype=2)+
  scale_x_log10()+
  scale_y_log10()+
  labs(
    x=expression(psi),
    y=expression(chi)
  )
```


## References
