---
title: "Testing the SMC on the Linear birth-death-sampling model"
author: "Aaron A. King"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
bibliography: phylopomp.bib
csl: jss.csl
params:
  prefix: "lbdp_smc"
---

```{r knitr-opts,child="knitr_setup.Rmd"}
```

Here, we simulate a genealogy under a linear birth-death-sampling process and compute its likelihood using the **phylopomp** SMC algorithm.
We compare the latter to the exact likelihood as computed using the formula of @Stadler2010.



```{r pkgs,results="hide",message=FALSE,cache=FALSE}
library(tidyverse)
library(pomp)
library(phylopomp)
stopifnot(packageVersion("phylopomp")>="0.0.11.0")
```

```{r simdat,results="hide"}
set.seed(257012390L)
playLBDP(lambda=2,mu=1,psi=1,times=3,n0=1) %>%
  getInfo(prune=TRUE) -> x
x %>% getInfo(compact=TRUE) %>% plot(points=TRUE)
x$tree %>%
  newick2df(time=3) -> dat
```

```{r smc_loglik,eval=FALSE}
dat %>%
  lbdp_pomp(lambda=2,mu=1,psi=1,method="gillespie") -> pogill

dat %>% 
  lbdp_pomp(lambda=2,mu=1,psi=1,method="euler",delta.t=0.001) -> poeuler

expand_grid(
  lambda=c(1.9,2.0,2.1),
  mu=c(0.9,1.0,1.1),
  psi=c(0.9,1.0,1.1),
  n0=1,
  Np=c(1e2,1e3,1e4,1e5,1e6),
  rep=1:10
) -> res

library(foreach)
library(iterators)
library(doParallel)
library(doRNG)
registerDoParallel()
registerDoRNG(558976750)

foreach (
  p=iter(res,"row"),
  .combine=bind_rows,
  .inorder=FALSE
) %dopar% {
  pogill %>%
    pfilter(
      Np=p$Np,
      params=unlist(p)
    ) %>%
    logLik() -> p$loglikGill
  
  pogill %>%
    pfilter(
      Np=p$Np,
      params=unlist(p)
    ) %>%
    logLik() -> p$loglikEuler
  
  dat %>%
    lbdp_exact(lambda=p$lambda,mu=p$mu,psi=p$psi) -> p$exact
  p
} -> res
```

```{r smc_loglik_eval,include=FALSE}
bake(file="smc_loglik(1).rds",{
  <<smc_loglik>>
}) -> res
```


```{r smc_loglik_plot,fig.dim=c(6,6),out.width="100%"}
library(reshape2)

res %>%
  select(-n0, -rep) %>%
  melt(id.vars=c("lambda","mu","psi","Np","exact")) %>%
  group_by(lambda,mu,psi,Np,exact,variable) %>%
  summarise(
    p25=quantile(value,probs=0.025)[[1]],
    p975=quantile(value,probs=0.975)[[1]],
    p50=quantile(value,probs=.5)[[1]]
  ) %>%
  ungroup() -> test

test %>%
  ggplot(aes(x=Np, y=p50, group=variable, color=factor(lambda))) +
  geom_point(aes(shape=variable), position=position_dodge(width=0.9))+
  geom_errorbar(aes(x=Np,ymax=p25,ymin=p975,linetype=variable),na.rm=TRUE,position="dodge") +
  geom_hline(aes(color=factor(lambda),yintercept=exact))+
  facet_grid(
    mu~psi,
    labeller=label_bquote(
      rows=mu==.(mu),
      cols=psi==.(psi)
    )
  )+
  labs(
    color=expression(lambda),
    x=expression(N[p]),
    y=expression(log~L)
  )+
  scale_x_log10(
    labels=function(x)aakmisc::scinot(x,simplify=TRUE)
  )+
  theme_bw(base_size=12)
```

Dots show the results of SMC computations using the indicated number of particles ($N_p$).
The horizontal lines are the exact values, as computed by @Stadler2010.

## Session Info

```{r session_info,echo=FALSE}
sessionInfo()
```

## References
