---
title: Likelihood slice for the SIRS model
author: A.A. King
bibliography: phylopomp.bib
nocite: |
  @King2024, @King2022
csl: jss.csl
vignette: >
  %\VignetteIndexEntry{SIRS computations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format:
  html:
    code-fold: show
    embed-resources: true
knitr:
  opts_chunk:
    progress: TRUE
    prompt: FALSE
    tidy: FALSE
    hightlight: TRUE
    strip.white: TRUE
    message: FALSE
    warning: FALSE
    error: TRUE
    echo: TRUE
    cache: TRUE
    cache.extra: rand_seed
    results: markup
---

```{r packages}
#| include: false
#| cache: false
library(tidyverse)
library(cowplot)
library(pomp)
library(phylopomp)
stopifnot(getRversion() >= "4.5")
stopifnot(packageVersion("pomp") >= "6.4")
stopifnot(packageVersion("phylopomp") >= "0.17.1")
theme_set(theme_bw(base_family = "serif"))
options(
  dplyr.summarise.inform = FALSE,
  pomp_archive_dir = "results/sirs"
)
set.seed(1159254136)
```

The following codes simulate a realization of the genealogy process induced by the SIRS model.

```{r simulate}
runSIRS(
  t0 = 0,
  time = 40,
  Beta = 4,
  gamma = 2,
  psi = 1,
  omega = 1,
  pop = 100,
  S0 = 0.97,
  I0 = 0.03,
  R0 = 0
) |>
  bake(
    file = "sirs1.rds",
    seed = 328168304L
  ) -> sim
```

The following constructs a **pomp** object that enables solving the filter equation for the simulated genealogy.

```{r process_data}
sim |>
  sir_pomp(
    Beta = 4,
    gamma = 2,
    psi = 1,
    omega = 1,
    pop = 100,
    S0 = 0.97,
    I0 = 0.03,
    R0 = 0
  ) -> sirs_pomp
```

The following solves the filter equation at each point of a grid.

```{r pfilter1}
#| code-fold: show
{
  library(doParallel)
  library(doRNG)
  library(iterators)
  registerDoParallel()
  registerDoRNG(728604304)

  expand_grid(
    Beta = 4,
    gamma = 2,
    psi = 1,
    omega = seq(0.5, 2, by = 0.05),
    pop = 100,
    S0 = 97,
    I0 = 3,
    R0 = 0,
    t0 = 0,
    rep = 1:16,
    Np = 5000,
  ) -> params

  foreach(
    p = iter(params, "row"),
    .inorder = FALSE,
    .combine = bind_rows
  ) %dopar%
    {
      sirs_pomp |>
        pfilter(Np = p$Np, params = p) |>
        logLik() |>
        bind_cols(p, logLik = _)
    } -> params
} |>
  bake(
    file = "sirs2.rds",
    dependson = sim
  ) -> params
```

The above computation took `r signif(attr(params,"system.time")[3]/60,3)`&nbsp;min.
The results are shown below.

```{r sirs1_plot}
#| out.width: 100%
#| fig.height: 8
#| fig.width: 8
sim |> plot() -> pl1

params |>
  group_by(Beta, gamma, psi, omega, S0, I0, R0, Np) |>
  reframe(
    name = c("logLik", "logLik_se", "logLik_ess"),
    value = logmeanexp(logLik, se = TRUE, ess = TRUE)
  ) |>
  ungroup() |>
  pivot_wider() |>
  mutate(
    y = logLik,
    ymax = logLik + 2 * logLik_se,
    ymin = logLik - 2 * logLik_se
  ) |>
  filter(round(Beta, 2) == 4) |>
  filter(logLik > max(logLik) - 16) |>
  ggplot(aes(x = omega, group = Np, y = y)) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "dodge") +
  labs(
    color = character(0),
    y = "log likelihood",
    x = expression(omega)
  ) +
  theme(
    legend.position = c(0.5, 0.27)
  ) -> pl2

plot_grid(
  A = pl1,
  B = pl2,
  labels = "AUTO",
  ncol = 1,
  rel_heights = c(1, 1)
)
```

## References
